A Wait-Free Queue with Polylogarithmic Step Complexity

### 2023-06-13 presentation/podc-presentation.tex
finished first draft of presentation

\only<1>{\input{MSQueue-48to50.pdf_t}}
\only<2>{\input{MSQueue-48to51.pdf_t}}
\only<3>{\input{MSQueue-48to52.pdf_t}}
\only<4>{\input{MSQueue-48-50to53.pdf_t}}
\only<5>{\input{MSQueue-50to54.pdf_t}}
\end{frame}

@@

\only<1>{\input{CASretry-49to50.pdf_t}}
\only<2>{\input{CASretry-49to51.pdf_t}}
\only<3->{\input{CASretry-49to52.pdf_t}}

@@

\begin{block}{New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\log^2 p + \log q)$ steps per \op{Dequeue}
\item wait-free
\item uses CAS on reasonable-size words
\item bounded space version: $O(\log p\ \log(p+q))$ amortized steps per operation
(relies on safe GC)
\end{itemize}
\end{block}

@@

\begin{frame}{Ordering Tree}
\visible<5>{\hfill Use ordering in root as linearization\hfill}
\medskip

\only<1>{\input{OrderTree-50.pdf_t}}%
\only<2>{\input{OrderTree-50to51.pdf_t}}%
\only<3>{\input{OrderTree-50to52.pdf_t}}%
\only<4->{\input{OrderTree-50to53.pdf_t}}

@@

Double Refresh:\\
If your CAS on $v$ fails twice, then\\
another process has propagated your operation to $v$.

@@

@@

\begin{itemize}
\item
Practical implementation\\
(perhaps slow path of fast path slow path method)
\item
Extend technique to other data structures\\
 (stacks and deques are recently done)
 \item
Close gap between\\
$\Omega(\log \log p)$ lower bound \hfill {\small [Attiya Fouren 2017]}\\
$O(\log^2 p + \log q)$ upper bound \hfill {\small [this work]}

>>>>

\only<1>{\input{MSQueue-1}}% 48-50
\only<2>{\input{MSQueue-2}}% 48-51
\only<3>{\input{MSQueue-3}}% 48-52
\only<4>{\input{MSQueue-4}}% 48,50-53
\only<5>{\input{MSQueue-5}}% 50-54
\end{frame}

@@

\only<1>{\input{CASretry-1}}% 49-50
\only<2>{\input{CASretry-2}}% 49-52
\only<3->{\input{CASretry-3}}% 49-52

@@

\begin{frame}{Breaking Linear-Time Bottleneck}

\begin{block}{Our New Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\log^2 p + \log q)$ steps per \op{Dequeue}
\item wait-free
\item uses CAS on reasonable-size words
\item bounded space version: $O(\log p\ \log(p+q))$ amortized steps per operation
(relies on safe GC)
\end{itemize}
\end{block}

@@

\begin{frame}{Ordering Tree}
\visible<5>{\hfill Use ordering in root as linearization\hfill}

\hspace*{-7mm}
\only<1>{\input{OrderTree-1}}% 50
\only<2>{\input{OrderTree-2}}% 50-51
\only<3>{\input{OrderTree-3}}% 49-52
\only<4->{\input{OrderTree-4}}% 50-53

@@

\Red{Double Refresh}:\\
If your CAS on $v$ fails twice, then\\
another process has propagated your operation to $v$.

@@

\begin{frame}{Requirements for Implicit Representation}

% 60% size; others are 80%
\hspace*{7mm}
\only<1>{\input{OrderTree-5}}% 50-53
\only<2>{\input{OrderTree-6}}% 48,50-53
\only<3>{\input{OrderTree-7}}% 50-54
\only<4>{\input{OrderTree-8}}% 50-55
\only<5>{\input{OrderTree-9}}% 50-56

Must support following in polylog time
\begin{itemize}
\pause
\item Refresh: promote batch of ops from children to parent
\pause
\item Find my \op{Dequeue} in root
\pause
\item Check if \op{Dequeue} returns null, or otherwise\\
determine rank of \op{Dequeue} among non-null \op{Dequeues}
\pause
\item Find \op{Enqueue} of given rank (and its argument)
\end{itemize}

\begin{frame}{Implicit Representation:  Blocks}

\only<1>{\input{OrderTree-5}}% 50-53
\only<2->{\input{OrderTree-10}}% 47,50-53

\vfill

\only<1-2>{\input{Blocks-0}}% none
\only<3>{\input{Blocks-1}}% 50
\only<4>{\input{Blocks-2}}% 49-50
\only<5>{\input{Blocks-3}}% 47-51
\only<6>{\input{Blocks-4}}% 46-51

\begin{frame}{Adding a Block for a Refresh}
\input{OrderTree-6}% 48,50-53

\vfill

\input{Blocks-5}% 46,48-50,52
\end{frame}

\begin{frame}{Tracing a \op{Dequeue} to the Root}
\input{OrderTree-7}% 50-54

\vfill

\input{Blocks-6}% 46,48-51,53,54
\end{frame}

\begin{frame}{Check if \op{Dequeue} Returns Null}
\input{OrderTree-7}% 50-54

\vfill

\input{Blocks-7}
\end{frame}

\begin{frame}{Rank of \op{Dequeue} Among Non-Null \op{Dequeues}}
\input{OrderTree-8}%

\vfill

\input{Blocks-8}
\end{frame}

\begin{frame}{Find \op{Enqueue} of Given Rank}
\input{OrderTree-9}

\vfill

\input{Blocks-9}
\end{frame}

\begin{block}{New Wait-Free Queue}
\begin{itemize}
\item $O(\log p)$ steps per \op{Enqueue}
\item $O(\mbox{log}^2 p + \log q)$ steps per \op{Dequeue}
\item<2> Unbounded space
\end{itemize}
\end{block}

$p = \#$ processes\\
$q = \#$ elements in queue
\end{frame}

\begin{frame}{Bounding Space}

\begin{itemize}
\item Replace array of blocks with a red-black tree of blocks
\item Periodically split RBT and discard obsolete blocks
\item Processes  help one another to ensure blocks are obsolete
\end{itemize}

\begin{block}{Bounded-Space Queue}
\begin{itemize}
\item Amortized $O(\log p\ \log(p+q))$ steps per operation
\item $O(pq+p^3\log p)$ space
\item Still wait-free
\end{itemize}
\end{block}

@@

\begin{itemize}
\item
Practical implementation\\
(perhaps slow path of fast path slow path method)
\item
Extend technique to other data structures\\
 (stacks and deques are recently done)
 \item
Close gap between\\
$\Omega(\log \log p)$ lower bound \hfill {\small [Attiya Fouren 2017]}\\
$O(\mbox{log}^2 p + \log q)$ upper bound \hfill {\small [this work]}
